Formalization of Forcing in Isabelle/ZF \thanks{Supported by Secyt-UNC project 33620180100465CB and Conicet.}

### 2020-04-12 IJCAR2020/proper_extension.tex
mas precisión sobre consistencia

Even when the axioms of $\ZFC$ are proved in the generic extension,
one cannot claim that the magic of forcing has taken place unless one
is able to provide some \emph{proper} extension with the \emph{same
ordinals}. After all, one is assuming from starters a model $M$ of $\ZFC$,
and in some trivial cases $M[G]$ might end up to be exactly $M$; this
is where \emph{proper} enters the stage. But, for instance, in the
presence of large cardinals, a model $M'\supsetneq M$ might be an
end-extension of $M$ ---this is were we ask the two models to have the
same ordinals, the same \emph{height}.

>>>>

Even when the axioms of $\ZFC$ are proved in the generic extension,
one cannot claim that the magic of forcing has taken place unless one
is able to provide some \emph{proper} extension with the \emph{same
ordinals}. After all, one is assuming from starters a model $M$ of $\ZFC$,
and in some trivial cases $M[G]$ might end up to be exactly $M$; this
is where \emph{proper} enters the stage. But, for instance, in the
presence of large cardinals, a model $M'\supsetneq M$ might be an
end-extension of $M$ ---this is where we ask the two models to have the
same ordinals, the same \emph{height}.

### 2020-04-12 IJCAR2020/related.tex
mas precisión sobre consistencia

\begin{itemize}
\item The consistency strength of Lean seems to require infinitely
  many inaccessibles. More precisely, Carneiro proved the  consistency
  of the CiC foundations used by Lean from $\ZFC$ plus the assumption
  of the existence of countably infinitely many inaccessible cardinals
  (as claimed by Hales \cite{hales-math-defs-formally} at the Big
  Proof 2019 workshop).%
  \footnote{We do not have a direct reference to Carneiro's proof.}

>>>>

\begin{itemize}
\item The consistency strength of Lean requires infinitely
  many inaccessibles. More precisely, let Lean$_n$ be the theory of
  CiC foundations of Lean restricted to $n$ type universes.  Carneiro
  \cite{carneiro-ms-thesis}, proved the consistency of Lean$_n$ from
  $\ZFC$ plus the existence of $n$ inaccessible
  cardinals. It is also reported in \cite{carneiro-ms-thesis} that
  Werner's results in \cite{10.5555/645869.668660} can be adapted to
  show that Lean$_{n+2}$ proves the consistency of the latter theory.
